# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Start by unzipping the compressed file and reading the data into R.

```{r}
activity <- read.csv(unz(paste(getwd(),"/activity.zip", sep=""), "activity.csv"), header=TRUE, sep=",")
```

## What is mean total number of steps taken per day?

```{r meandaily, echo=FALSE}
hist(activity$steps)
meen <- mean(activity$steps, na.rm=TRUE)
meediin <- median(activity$steps, na.rm=TRUE)
```

The mean steps per day from this data set is `r meen`.
The median number of steps per day is `r meediin`.

## What is the average daily activity pattern?

```{r avgdaily, echo=FALSE}
avg_steps_per_interval <- with(activity, tapply(steps, interval, mean, na.rm=TRUE))
plot(avg_steps_per_interval, type="l")
macks <- max(avg_steps_per_interval, na.rm=TRUE)
macks_int <- which(avg_steps_per_interval==max(avg_steps_per_interval))
```

This individual averaged `r macks` steps during the `r macks_int`th interval.

## Imputing missing values

```{r missing,echo=FALSE}
missing_no <- nrow(activity) - sum(complete.cases(activity))
```

Unfortunately, there are `r missing_no` observations in the data that are missing the number of steps taken.

We compensate by substituting the average number of steps for a given five-minute interval across all days whenever data for an individual day does not exist.

```{r replace, echo=FALSE}
activity2 <- activity
activity2$average <- rep(avg_steps_per_interval * -1)
activity2$steps2 <- pmax(activity2$steps, activity2$average, na.rm=TRUE)
activity2$steps <- abs(activity2$steps2)
hist(activity2$steps)
meen2 <- mean(activity2$steps, na.rm=TRUE)
meediin2 <- median(activity2$steps)
```

The new mean steps per day from this data set is `r meen2`.
The new median number of steps per day is `r meediin2`.

## Are there differences in activity patterns between weekdays and weekends?

```{r dayofweek, echo=FALSE}
activity2$date <- as.Date(activity2$date)
activity2$dayofweek <- weekdays(activity2$date)
weekday_data <- subset(activity2, dayofweek!="Sunday" & dayofweek!="Saturday")
weekend_data <- subset(activity2, dayofweek=="Sunday" | dayofweek=="Saturday")
```

```{r avgdaily2, echo=FALSE}
avg_steps_per_interval_D <- with(weekday_data, tapply(steps, interval, mean, na.rm=TRUE))
plot(avg_steps_per_interval_D, type="l")
avg_steps_per_interval_E <- with(weekend_data, tapply(steps, interval, mean, na.rm=TRUE))
plot(avg_steps_per_interval_E, type="l")
```
